---
title: "Data_Transformations"
author: "Jessica Bullington"
date: "4/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### This document explores transformations of my Ala Wai Canal V. vulnificus dataset in the hopes of identifying appropriate transformations for all environmental predictors and response variable (vvhA copies per mL).

### Load the data.
```{r}
rm(list=ls()) # clear workspace
data = read.csv("AlaWai_DiscreteBottle.csv") # first drag/drop in repo on github website
data = subset(data, !is.na(data$vvhA)) # keep only rows with observations for vvhA
data$vvhA = as.numeric(data$vvhA) # change the class type to numeric
```


### Explore the data visually
```{r}
# Histograms of all numerical data
for (i in 5:ncol(data)){ # for each column, make two plots
  hist(data[,i], xlab=colnames(data)[i], main = paste("Histogram", colnames(data)[i])) 
  qqnorm(data[,i], xlab=colnames(data)[i], main = paste("SW=", shapiro.test(data[,i])$p.value), pch = 1, frame = FALSE) # qq-plot with SW p-value
  qqline(data[,i], col = "steelblue", lwd = 2) # perfectly normal line
}
```

#### Conclusion: Temperature doesn't need a transformation.

```{r}
# Histograms of all numerical data
for (i in 5:ncol(data)){ # for each column, make two plots
  hist(log10(data[,i]+1), xlab=colnames(data)[i], main = paste("Log10 +1 Histogram", colnames(data)[i])) 
  qqnorm(log10(data[,i]+1), xlab=colnames(data)[i], main = paste("SW=", shapiro.test(log10(data[,i]+1))$p.value), pch = 1, frame = FALSE) # qq-plot with SW p-value
  qqline(log10(data[,i]+1), col = "steelblue", lwd = 2) # perfectly normal line
}
```
#### Conclusion: Log10+1 transform works for TotalN, Turbidity, POC, DOC.

```{r}
# Histograms of all numerical data
for (i in 5:ncol(data)){ # for each column, make two plots
  hist(data[,i]^(1/3), xlab=colnames(data)[i], main = paste("Cubrt Histogram", colnames(data)[i])) 
  qqnorm(data[,i]^(1/3), xlab=colnames(data)[i], main = paste("SW=", shapiro.test(data[,i]^(1/3))$p.value), pch = 1, frame = FALSE) # qq-plot with SW p-value
  qqline(data[,i]^(1/3), col = "steelblue", lwd = 2) # perfectly normal line
}
```
#### Conclusion: Cube root works for TotalP, Si, Chl. (Note: squareroot did not and fourth root was slightly better for Chl.) N+N, FCM, and Depth still skewed!


### Check out transformations of vvhA in more detail

#### Basic transformations
```{r}
# Define a set of transformations to test
n = data$vvhA # nothing
log = log10(data$vvhA + 1) # add 1 otherwise undefined
ln = log(data$vvhA + 1)
sqrt = sqrt(data$vvhA) # reduce powers since right skewed
cbrt = (data$vvhA)^(1/3)
fort = (data$vvhA)^(1/4)
ffrt = (data$vvhA)^(1/5)
sxrt = (data$vvhA)^(1/6)
  
# Combine the output in a data frame
tran = data.frame(n, log, ln, sqrt, cbrt, fort, ffrt, sxrt)

# Visually inspect each transformation with a histogram and quantile-quantile plot
# Plot Shapiro-Wilk test result in qq-plot title

par(mfrow=c(4,ncol(tran)/2)) # combine the following plots into one grid
for (i in 1:ncol(tran)){ # for each column, make two plots
  hist(tran[,i], xlab=colnames(tran)[i], main = paste("Histogram", colnames(tran)[i], "(vvhA)")) # histogram
  qqnorm(tran[,i], xlab=colnames(tran)[i], main = paste("SW=", shapiro.test(tran[,i])$p.value), pch = 1, frame = FALSE) # qq-plot with SW p-value
  qqline(tran[,i], col = "steelblue", lwd = 2) # perfectly normal line
}

# None of these look great but 1/4 is the closest
# Let's see what the optimum would be to maximize the SW p-value

shapiro.pvalue <- function(a, x){ # define the function to be optimized
  shapiro.test(x^a)$p.value 
}
optimize(shapiro.pvalue, lower=0.01, upper=1, x=data$vvhA, maximum=TRUE) 

# In conclusion, the optimal power function would be 0.2280655 which gives a SW p-value of 0.004184215 so technically we're still not in the clear.

```

#### Advanced transformations: IHS and BoxCox from https://stats.stackexchange.com/questions/1444/how-should-i-transform-non-negative-data-including-zeros

##### Also consider more complex modeling techniques: GLMs with poisson distribution, zero-inflated poisson, mixture models, and quantile regression. Let's deal with those later and try transforming first.

##### Inverse hyperbolic sine (IHS) transformation
```{r}
# IHS is used by economists when there is a jump to people with zero income
# https://stats.stackexchange.com/questions/78929/how-can-i-estimate-theta-for-the-inverse-hyperbolic-sine-transformation
# https://stats.stackexchange.com/questions/78929/how-can-i-estimate-theta-for-the-inverse-hyperbolic-sine-transformation

ihs <- function(x, theta) { # define the ISH function
  #log(theta*x + sqrt((theta^2)*(x^2) + 1))/theta # this should be equivalent to the following
  (1/theta)*asinh(theta*x)
}

shapiro.pvalue <- function(theta, x){ # define the function to be optimized
  x <- ihs(x, theta)
  shapiro.test(x)$p.value # can separate
}

optimize(shapiro.pvalue, lower=0.001, upper=50, x=data$vvhA, maximum=TRUE) 

# In conclusion, the optimal theta value would be 2.435218 which gives a SW p-value of 1.269239e-05 so this is worse than power transform.

```


#### Box Cox transformation power selection
```{r}
# The Box Cox transformation works on the residuals of a linear model. It can improve normality and homoscedasticity (equal variance among different groups).
# https://www.r-bloggers.com/how-to-detect-heteroscedasticity-and-rectify-it/
# https://rpubs.com/bskc/288328



```


## Salinity
```{r}
# Define a set of transformations to test
n = data$Salinity # nothing
log = log10(100 - data$Salinity) # add 1 otherwise undefined
ln = log(100 - data$Salinity)
sq = data$Salinity^2 # reduce powers since right skewed
cb = data$Salinity^3
fo = data$Salinity^4
ff = data$Salinity^5
scale_fo = (data$Salinity^4)/10000
  
# Combine the output in a data frame
tran = data.frame(n, log, ln, sq, cb, fo, scale_fo, ff)

# Visually inspect each transformation with a histogram and quantile-quantile plot
# Plot Shapiro-Wilk test result in qq-plot title

par(mfrow=c(4,ncol(tran)/2)) # combine the following plots into one grid
for (i in 1:ncol(tran)){ # for each column, make two plots
  hist(tran[,i], xlab=colnames(tran)[i], main = paste("Histogram", colnames(tran)[i], "(Salinity)")) # histogram
  qqnorm(tran[,i], xlab=colnames(tran)[i], main = paste("SW=", shapiro.test(tran[,i])$p.value), pch = 1, frame = FALSE) # qq-plot with SW p-value
  qqline(tran[,i], col = "steelblue", lwd = 2) # perfectly normal line
}

```





