---
title: "Data_Transformations"
author: "Jessica Bullington"
date: "4/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### This document explores transformations of my Ala Wai Canal V. vulnificus dataset in the hopes of identifying appropriate transformations for all environmental predictors and response variable (vvhA copies per mL).

### Load the data.
```{r}
rm(list=ls()) # clear workspace
data = read.csv("AlaWai_DiscreteBottle.csv") # first drag/drop in repo on github website
data = subset(data, !is.na(data$vvhA)) # keep only rows with observations for vvhA
data$vvhA = as.numeric(data$vvhA) # change the class type to numeric
```

### Check out transformations of vvhA

#### Basic transformations
```{r}
# Define a set of transformations to test
n = data$vvhA # nothing
log = log10(data$vvhA + 1) # add 1 otherwise undefined
ln = log(data$vvhA + 1)
sqrt = sqrt(data$vvhA) # reduce powers since right skewed
cbrt = (data$vvhA)^(1/3)
fort = (data$vvhA)^(1/4)
ffrt = (data$vvhA)^(1/5)
sxrt = (data$vvhA)^(1/6)
  
# Combine the output in a data frame
tran = data.frame(n, log, ln, sqrt, cbrt, fort, ffrt, sxrt)

# Visually inspect each transformation with a histogram and quantile-quantile plot
# Plot Shapiro-Wilk test result in qq-plot title

par(mfrow=c(4,ncol(tran)/2)) # combine the following plots into one grid
for (i in 1:ncol(tran)){ # for each column, make two plots
  hist(tran[,i], xlab=colnames(tran)[i], main = paste("Histogram", colnames(tran)[i], "(vvhA)")) # histogram
  qqnorm(tran[,i], xlab=colnames(tran)[i], main = paste("SW=", shapiro.test(tran[,i])$p.value), pch = 1, frame = FALSE) # qq-plot with SW p-value
  qqline(tran[,i], col = "steelblue", lwd = 2) # perfectly normal line
}

# None of these look great but 1/4 is the closest
# Let's see what the optimum would be to maximize the SW p-value

shapiro.pvalue <- function(a, x){ # define the function to be optimized
  shapiro.test(x^a)$p.value 
}
optimize(shapiro.pvalue, lower=0.01, upper=1, x=data$vvhA, maximum=TRUE) 

# In conclusion, the optimal power function would be 0.2280655 which gives a SW p-value of 0.004184215 so technically we're still not in the clear

```

#### Advanced transformations
https://stats.stackexchange.com/questions/1444/how-should-i-transform-non-negative-data-including-zeros

##### Inverse hyperbolic sine (IHS) transformation
```{r}
data$vvhA.IHS = log10(data$vvhA + ((data$vvhA^2 + 1)^(1/2)))
hist(data$vvhA.IHS)
qqPlot(data$vvhA.IHS)
shapiro.test(data$vvhA.IHS)

ihs <- function(x) {
  y <- log(x + sqrt(x ^ 2 + 1))
  return(y)
}
```


#### Box Cox transformation power selection
```{r}
#https://www.r-bloggers.com/how-to-detect-heteroscedasticity-and-rectify-it/
#https://rpubs.com/bskc/288328
```








