---
title: "Model_Selection"
author: "Jessica Bullington"
date: "5/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load the data
```{r}
rm(list=ls()) # clear workspace
data = read.csv("AlaWai_DiscreteBottle_20200728.csv") # first drag/drop in repo on github website
#data = subset(data, !is.na(data$vvhA)) # keep only rows with observations for vvhA
data$vvhA = as.numeric(data$vvhA) # change the class type to numeric
data$vvhA_ND = as.numeric(data$vvhA_ND) # change the class type to numeric

```

### Transformations
```{r}
# make a new column for each transformation
data$tran.vvhA = (data$vvhA)^(1/4)
data$tran.vvhA_ND = log10(data$vvhA_ND)
data$tran.TN = log10(data$TotalN+1)
data$tran.TP = (data$TotalP)^(1/3)
data$tran.Si = (data$Si)^(1/3)
# N+N?
data$tran.DOC = log10(data$DOC+1)
data$tran.POC = log10(data$POC+1)
# FCM?
data$tran.Sal = (data$Salinity^4)/10000 # dividing by 10000 removes the scaling issue with lm
data$tran.Temp = data$Temperature
data$tran.Turb = log10(data$Turbidity+1)
data$tran.Chl = (data$Chlorophyll)^(1/4)
# Depth?
#write.csv(data, "AlaWaiDiscreteBottle_transforms_21May2020.csv")
```


### Look for covariance in predictors
```{r fig.width = 9, fig.height = 8}
pairs(~tran.TN + tran.TP + tran.Si + tran.DOC + tran.POC + tran.Sal + tran.Temp + tran.Turb + tran.Chl, data =data) # add N+N, FCM, Depth
```


### Model selection
```{r}
library(MuMIn)
# Subset the data to only the variables being tested
var = c("tran.vvhA_ND",  "tran.TN", "tran.TP", "tran.Si", "tran.DOC", "tran.POC", "tran.Sal", "tran.Temp", "tran.Turb", "tran.Chl", "Month", "Site")
dredge.data = data[var]
dredge.data = na.omit(dredge.data) # Note that dredge cannot handle missing values

# Fit a big model that includes all predictors
options(na.action = "na.fail")
big.model = lm(tran.vvhA_ND ~ tran.TN + tran.TP + tran.Si + tran.DOC + tran.POC + tran.Sal + tran.Temp + tran.Turb + tran.Chl, data =dredge.data)

# LRT
summary(big.model)

# All possible model combinations
big.dredge = dredge(big.model, extra = "R^2")

# Model selection based on AIC (rather than stepwise)
head(model.sel(big.dredge), n=20)

# Multimodel inference
big.avg = model.avg(big.dredge, delta < 4)
summary(big.avg)
confint(big.avg)
importance(big.avg)
# add interactions?
```

#### Conclusion: Site is not useful. Month, salinity, turbidity important based on LRT. Month, salinity, silicate, temperature, total phosphorous important based on AIC. However, salnitiy and temperautre correlate with the nutrients which might be problematic for AIC.

### Model selection with limited model
```{r fig.width = 9, fig.height = 8}
# remove correlating predictors
# remove site
# add season and habitat? ~ instead of month and site

pairs(~tran.Si + tran.DOC + tran.POC + tran.Sal + tran.Temp + tran.Turb, data =data) # add N+N, FCM, Depth

# Subset the data to only the variables being tested
# Need to run on same dataset to compare AIC above
# var = c("tran.vvhA", "tran.TN", "tran.TP", "tran.Si", "tran.DOC", "tran.POC", "tran.Sal", "tran.Temp", "tran.Turb", "tran.Chl", "Month", "Site")
# dredge.data = data[var]
# dredge.data = na.omit(dredge.data) # Note that dredge cannot handle missing values

# Fit a big model that includes all predictors
options(na.action = "na.fail") # no missin values
big.model = lm(tran.vvhA_ND ~ tran.Si + tran.DOC + tran.POC + tran.Sal + tran.Temp + tran.Turb, data =dredge.data)

# LRT
summary(big.model)

# All possible model combinations
big.dredge = dredge(big.model, extra = "R^2")

# Model selection based on AIC (rather than stepwise)
head(model.sel(big.dredge), n=20)

# Multimodel inference
big.avg = model.avg(big.dredge, delta < 4)
summary(big.avg)
confint(big.avg)
importance(big.avg)
```

#### Conclusion: Month, salinity, and temperature are important based on LRT. Month, salinity, silicate, and temperature are important based on AIC.

### Test interaction
```{r}
# Fit a big model that includes all predictors
options(na.action = "na.fail") # no missin values
big.model = lm(tran.vvhA ~ tran.Si + tran.DOC + tran.POC + tran.Sal + tran.Temp + tran.Turb + Month + tran.Sal:Month, data =dredge.data)

# LRT
summary(big.model)

# All possible model combinations
big.dredge = dredge(big.model, extra = "R^2")

# Model selection based on AIC (rather than stepwise)
head(model.sel(big.dredge), n=20)

# Multimodel inference
big.avg = model.avg(big.dredge, delta < 4)
summary(big.avg)
confint(big.avg)
importance(big.avg)
```

#### Conclusion: You can use interactions with dredge. Month has a significant interaction with salinity (not temp). Possibly run interactions on best model at end.

### Add fDOM
```{r fig.width = 9, fig.height = 8}
data$tran.Humic = data$MarineHumic.like^(1/3)
data$tran.Protein = data$Tyrosine.like^(1/3)
  
pairs(~tran.Si + tran.DOC + tran.POC + tran.Sal + tran.Temp + tran.Turb + tran.Humic + tran.Protein, data =data)

# Subset the data to only the variables being tested
var = c("tran.vvhA", "tran.TN", "tran.TP", "tran.Si", "tran.DOC", "tran.POC", "tran.Sal", "tran.Temp", "tran.Turb", "tran.Chl", "Month", "tran.Humic", "tran.Protein")
dredge.data = data[var]
dredge.data = na.omit(dredge.data) # Note that dredge cannot handle missing values

# Fit a big model that includes all predictors
options(na.action = "na.fail")
big.model = lm(tran.vvhA ~ tran.Si + tran.DOC + tran.POC + tran.Sal + tran.Temp + tran.Turb + tran.Chl + Month + tran.Humic + tran.Protein, data =dredge.data)

# LRT
summary(big.model)

# All possible model combinations
big.dredge = dredge(big.model, extra = "R^2")

# Model selection based on AIC (rather than stepwise)
head(model.sel(big.dredge), n=20)

# Multimodel inference
big.avg = model.avg(big.dredge, delta < 4)
summary(big.avg)
confint(big.avg)
importance(big.avg)
```